Benchmarking on Rigetti architecture

### 2022-07-04 txt/manuscript.tex
add pictures to manuscript
changes in proofs

\subsection{Theoretical bound}\label{sec:example_theoretical_probability}

\begin{theorem}
The optimal probability of correct discrimination between von Neumann
measurements $\PP_U$ and $\PP_{\Id}$ for $U = H \diag(1, e^{i \phi}) H^\dagger$,
where $\phi \in [0, 2\pi)$ is given by
\begin{equation}
p = \frac{1}{2} + \frac{|1 - e^{i \phi}  |}{4} . 
\end{equation}
\end{theorem}
\begin{proof}
	From Holevo-Helstrom theorem we obtain
	\begin{equation}
	p = \frac{1}{2} + \frac{1}{4} \| \PP_{U} - \PP_{\Id} \|_\diamond,
	\end{equation}
	therefore from Theorem 1 in \cite{puchala2018strategies} we have 
	\begin{equation}
	\|\PP_U - \PP_\Id\|_\diamond = \min_{E \in \diaguni_d} \|\Phi_{UE} - 
	\Phi_\Id\|_\diamond. 
	\end{equation}
	From Lemma~\ref{lemma:min-e-optimal} we show that 
	\begin{equation}
		\min_{E \in \diaguni_2} \|\Phi_{UE} - 
	\Phi_\Id\|_\diamond = \|\Phi_{U} - 
	\Phi_\Id\|_\diamond
	\end{equation} and hence, from the problem of discrimination of unitary channels, we receive 
	\begin{equation}
	p = \frac{1}{2} + \frac{1}{2} \sqrt{1 - \nu^2} = \frac{1}{2} + \frac{|1-e^{i \phi}|}{4},
	\end{equation} which completes the proof.
\end{proof}

\todo[inline]{change proof of proposition into proof of theorem}
\begin{lemma}\label{lemma:min-e-optimal}
 Let $U = H \diag(1, e^{i \phi}) H^\dagger$, $\phi \in [0, 2\pi)$ and	let 
 $\Phi_U$ and $\Phi_\Id$ be two unitary channels. The following equation holds 
	\begin{equation}
	\min_{E \in \diaguni_2} \|\Phi_{UE} - 
	\Phi_\Id\|_\diamond = \|\Phi_{U} - 
	\Phi_\Id\|_\diamond,
	\end{equation}

\end{lemma}
\begin{proof} We know that
$
	\| \Phi_U  - \Phi_{\1} \|_\diamond = 2 \sqrt{1-\nu^2},
$
	where $\nu = \min_{x \in W(U^\dagger)} |x|  $. For $U = H 
	\left(\begin{array}{cc}1&0\\0&e^{i \phi}\end{array}\right)  H^\dagger$ the readers briefly observe that $W(U^\dagger) $ 
	obtain that $\nu^2 = 1 - \frac{|1 - e^{i \phi} | }{2}$. 
	The condition $ 	\min_{E \in \diaguni_2} \|\Phi_{UE} - 
	\Phi_\Id\|_\diamond = \|\Phi_{U} - 
	\Phi_\Id\|_\diamond $ is equivalent to prove that
	\begin{equation}
	\max_{E \in \diaguni_2 } \nu_E = \nu = \frac{|1 + e^{i \phi} | }{2},
	\end{equation}
	where $\nu_E = \min_{x \in W(U^\dagger E)} |x|. $ We can prove that
	\begin{equation}
	\min_{\ket{x} \in \mathbb{C}^2  \proj{x} = 1} |\bra{x}U^\dagger\ket{x}| = 
	\min_{\rho \in \Omega_2} |\tr(U^\dagger\rho)|. 
	\end{equation}
For that, to calculate $\nu_E$ we obtain \begin{equation}
	\max_{E \in \diaguni_2 } \nu_E  = \max_{E \in \diaguni_2 }  \min_{\rho \in 
	\Omega_2} \left| \tr \rho U E  \right|
\end{equation}
For that, our task is reduce to show that
	\begin{equation}
	\forall E \in \diaguni_2 \,\, | \tr \rho U E | \le \nu. 
	\end{equation}
	Let us define $E = \left(\begin{array}{cc}E_0&0\\0&E_1\end{array}\right)  $ 
	and let us take $\rho = 
	\left(\begin{array}{cc}\frac{1}{2}&0\\0&\frac{1}{2}\end{array}\right) $. 
	From spectral theorem, let us note $U$ as
	\begin{equation}
	U= \lambda_1 \ketbra{x_1}{x_1} + \lambda_2 \ketbra{x_1}{x_2}, 
	\end{equation}
	where  for eigenvector $\lambda_1 = e^{i \phi}$ the eigenvector is 
	of the form $\ket{x_1} = 
	\left[\begin{array}{c}\frac{1}{\sqrt{2}}\\-\frac{1}{\sqrt{2}}\end{array}\right]$,
	 whereas for  $\lambda_2 = 1$ we have $\ket{x_2} = 
	\left[\begin{array}{c}\frac{1}{\sqrt{2}}\\\frac{1}{\sqrt{2}}\end{array}\right]$.
	Then, we have 
\begin{equation}
\begin{split}
& \forall E \in \diaguni_2 \,\,\, | \tr \rho U E | = \frac{1}{2}  \left| \tr 
H \diag(1, e^{i\phi}) H^\dagger E \right| =  \\ &
\frac{1}{2} \left| \tr\left((  e^{i \phi} \proj{x_1} + 1 \proj{x_2} ) E \right) 
\right|  = 
\frac{1}{2} \left| e^{i \phi}  \bra{x_1} E \ket{x_1} + \bra{x_2} E \ket{x_2} 
\right| = \\& 
\frac{1}{2} \left| \frac{E_0 + E_1}{2} + e^{i \phi } \frac{E_0+E_1}{2} \right| 
= 
\frac{\left| 1+ e^{i \phi } \right|}{2} \left| \frac{E_0 + E_1}{2} \right| \le 
\nu, 
	\end{split}
\end{equation}
	which completes the proof.
\end{proof}

@@

%The schematic representation of theoretical 
%setup is shown in Fig.~\ref{fig:theoretical}.

@@

%\begin{figure}[h!]
%	\centering 
%	\includegraphics[width=0.75\linewidth, angle=-90]{rys-postsellection.jpg} 
%	
%	\caption{ A schematic representation of the setup for distinguishing
%		measurements using postselection. 
%	}\label{fig:postsellection}
%\end{figure}
\begin{enumerate}
\item We prepare input state $\ket{\psi} = \frac{1}{\sqrt{2} } | \Id_2 \rangle 
\rangle. $
\item We prepare one of two unitary channel $\Phi_{U} $ or $\Phi_{\1}$. 
\item We implement unitary $V_0 $ or $ V_1$.
\item We prepare the measurement $\Delta$ in standard basis (already exists on 
Rigetti architecture) on each qubits.
\item We calculate the probability of correct discrimination in the following 
way
\begin{equation}
p = ...
\end{equation}
%\todo[inline]{grubsza kmina, musze Ci to na zywo opowiedziec xd}
\end{enumerate}
\end{scheme}

\begin{scheme}(By using controlled unitary)
%\begin{figure}[h!]
%	\centering 
%	\includegraphics[width=0.75\linewidth, angle=-90]{rys-controlled.jpg} 
%	
%	\caption{ A schematic representation of the setup for distinguishing
%		measurements using controlled unitary gate. 
%	}\label{fig:controlled}
%\end{figure}
\begin{enumerate}
\item We prepare input state $\ket{\psi} = \frac{1}{\sqrt{2} } | \Id_2 \rangle 
\rangle. $
\item We prepare one of two unitary channel $\Phi_{U} $ or $\Phi_{\1}$. 
\item We implement controlled unitary $V_0 \oplus V_1$.
\item We prepare the measurement $\Delta$ in standard basis (already exist on 
Rigetti architecture) on each qubits.
\item We calculate the probability of correct discrimination in the following 
way
\begin{equation}
p = \frac{|j=0 \wedge \Phi_{U} | + |j=1 \wedge \Phi_{\1}|}{\text{all 
cases}}
\end{equation}
\end{enumerate}
\end{scheme}

>>>>

\subsection{Theoretical probability}\label{sec:example_theoretical_probability}
In this section we calculate the theoretical probability of correct discrimination between two von Neumann measurements $\PP_\1$ and $\PP_U$ for $U = H \diag(1, e^{i \phi}) H^\dagger$, where  $\phi \in [0, 2\pi)$. To do that, we will present an auxiliary lemma.   
\begin{lemma}\label{lemma:min-e-optimal}
	Let $U = H \diag(1, e^{i \phi}) H^\dagger$, $\phi \in [0, 2\pi)$ and	let 
	$\Phi_U$ and $\Phi_\Id$ be two unitary channels. The following equation holds 
	\begin{equation}
	\min_{E \in \diaguni_2} \|\Phi_{UE} - 
	\Phi_\Id\|_\diamond = \|\Phi_{U} - 
	\Phi_\Id\|_\diamond,
	\end{equation}

\end{lemma}
\begin{proof} Recall that the distance between two unitary channels is given by
	$
	\| \Phi_U  - \Phi_{\1} \|_\diamond = 2 \sqrt{1-\nu^2},
	$
	where $\nu_U = \min_{x \in W(U^\dagger)} |x|  $ for any $U \in \mathcal{U}_d$.

	For $U = H 
	\left(\begin{array}{cc}1&0\\0&e^{i \phi}\end{array}\right)  H^\dagger$ the readers briefly observe that  $\nu^2 = 1 - \frac{|1 - e^{i \phi} | }{4}$. So, 
	\begin{equation}
	\|  \Phi_U  - \Phi_{\1} \|_\diamond = \frac{| 1 - e^{i \phi} |}{2}. 
	\end{equation} 
		It implies that it is enough to prove  \begin{equation}
		\min_{E \in \diaguni_2} \|\Phi_{UE} - 
		\Phi_\Id\|_\diamond  = \frac{| 1 - e^{i \phi} |}{2}.
		\end{equation}

		It implies that we can prove  equivalently  the following condition
		\begin{equation}
		\max_{E \in \diaguni_2 } \nu_{UE} = \nu_U
		\end{equation}
	It implies that it is enough to prove 
	\begin{equation}
	\max_{E \in \diaguni_2 } \nu_E = \nu = \frac{|1 + e^{i \phi} | }{2},
	\end{equation}
	where $\nu_E = \min_{x \in W(U^\dagger E)} |x|. $

	We can prove that
	\begin{equation}
	\min_{\ket{x} \in \mathbb{C}^2:   \proj{x} = 1} |\bra{x}U^\dagger\ket{x}| = 
	\min_{\rho \in \Omega_2} |\tr(U^\dagger\rho)|. 
	\end{equation}
	For that we obtain \begin{equation}
	\max_{E \in \diaguni_2 } \nu_{UE}  = \max_{E \in \diaguni_2 }  \min_{\rho \in 
		\Omega_2} \left| \tr \rho U E  \right|
	\end{equation}
	For that, our task is reduce to show that
	\begin{equation}
	\forall E \in \diaguni_2 \,\, | \tr \rho U E | \le \nu. 
	\end{equation}
	Let us define $E = \left(\begin{array}{cc}E_0&0\\0&E_1\end{array}\right)  $ 
	and let us take $\rho = 
	\left(\begin{array}{cc}\frac{1}{2}&0\\0&\frac{1}{2}\end{array}\right) $. 
	From spectral theorem, let us note $U$ as
	\begin{equation}
	U= \lambda_1 \ketbra{x_1}{x_1} + \lambda_2 \ketbra{x_1}{x_2}, 
	\end{equation}
	where  for eigenvector $\lambda_1 = e^{i \phi}$ the eigenvector is 
	of the form $\ket{x_1} = 
	\left[\begin{array}{c}\frac{1}{\sqrt{2}}\\-\frac{1}{\sqrt{2}}\end{array}\right]$,
	whereas for  $\lambda_2 = 1$ we have $\ket{x_2} = 
	\left[\begin{array}{c}\frac{1}{\sqrt{2}}\\\frac{1}{\sqrt{2}}\end{array}\right]$.
	Then, we have 
	\begin{equation}
	\begin{split}
	& \forall E \in \diaguni_2 \,\,\, | \tr \rho U E | = \frac{1}{2}  \left| \tr 
	H \diag(1, e^{i\phi}) H^\dagger E \right| =  \\ &
	\frac{1}{2} \left| \tr\left((  e^{i \phi} \proj{x_1} + 1 \proj{x_2} ) E \right) 
	\right|  = 
	\frac{1}{2} \left| e^{i \phi}  \bra{x_1} E \ket{x_1} + \bra{x_2} E \ket{x_2} 
	\right| = \\& 
	\frac{1}{2} \left| \frac{E_0 + E_1}{2} + e^{i \phi } \frac{E_0+E_1}{2} \right| 
	= 
	\frac{\left| 1+ e^{i \phi } \right|}{2} \left| \frac{E_0 + E_1}{2} \right| \le 
	\nu, 
	\end{split}
	\end{equation}
	which completes the proof.
\end{proof}
\begin{theorem}
The optimal probability of correct discrimination between von Neumann
measurements $\PP_U$ and $\PP_{\Id}$ for $U = H \diag(1, e^{i \phi}) H^\dagger$,
where $\phi \in [0, 2\pi)$ is given by
\begin{equation}
p = \frac{1}{2} + \frac{|1 - e^{i \phi}  |}{4} . 
\end{equation}
\end{theorem}
\begin{proof}
	From Holevo-Helstrom theorem we obtain
	\begin{equation}
	p = \frac{1}{2} + \frac{1}{4} \| \PP_{U} - \PP_{\Id} \|_\diamond,
	\end{equation}
	therefore from Theorem 1 in \cite{puchala2018strategies} we have 
	\begin{equation}
	\|\PP_U - \PP_\Id\|_\diamond = \min_{E \in \diaguni_d} \|\Phi_{UE} - 
	\Phi_\Id\|_\diamond. 
	\end{equation}
	From Lemma~\ref{lemma:min-e-optimal} we show that 
	\begin{equation}
		\min_{E \in \diaguni_2} \|\Phi_{UE} - 
	\Phi_\Id\|_\diamond = \|\Phi_{U} - 
	\Phi_\Id\|_\diamond
	\end{equation} and hence, from the problem of discrimination of unitary channels, we receive 
	\begin{equation}
	p = \frac{1}{2} + \frac{1}{2} \sqrt{1 - \nu^2} = \frac{1}{2} + \frac{|1-e^{i \phi}|}{4},
	\end{equation} which completes the proof.
\end{proof}

\todo[inline]{change proof of proposition into proof of theorem}

@@

The schematic representation of theoretical setup is shown in the following 
figure.

\begin{figure}[h!]
\centering
\includegraphics[scale=1.2]{pics/theoretical_scheme}
\caption{Theoretical discrimination scheme}
\label{fig:theoretical_scheme}
\end{figure}

@@

\begin{figure}[h!]
\centering 
\includegraphics[scale=1.2]{pics/postselection}

\caption{ A schematic representation of the setup for distinguishing
	measurements using postselection. 
}\label{fig:postsellection}
\end{figure}
\begin{enumerate}
\item We prepare input state $\ket{\psi} = \frac{1}{\sqrt{2} } | \Id_2 \rangle 
\rangle. $
\item We prepare one of two unitary channel $\Phi_{U} $ or $\Phi_{\1}$. 
\item We implement unitary $V_0 $ or $ V_1$.
\item We prepare the measurement $\Delta$ in standard basis (already exists on 
Rigetti architecture) on each qubits.
\item We calculate the probability of correct discrimination in the following 
way
\begin{equation}
p = ...
\end{equation}
%\todo[inline]{grubsza kmina, musze Ci to na zywo opowiedziec xd}
\end{enumerate}
\end{scheme}

\begin{scheme}(By using controlled unitary)
\begin{figure}[h!]
\centering 
\includegraphics[scale=1.2]{pics/controlled_unitary}

\caption{ A schematic representation of the setup for distinguishing
	measurements using controlled unitary gate. 
}\label{fig:controlled}
\end{figure}
\begin{enumerate}
\item We prepare input state $\ket{\psi} = \frac{1}{\sqrt{2} } | \Id_2 \rangle 
\rangle. $
\item We prepare one of two unitary channel $\Phi_{U} $ or $\Phi_{\1}$. 
\item We implement controlled unitary $V_0 \oplus V_1$.
\item We prepare the measurement $\Delta$ in standard basis (already exist on 
Rigetti architecture) on each qubits.
\item We calculate the probability of correct discrimination in the following 
way
\begin{equation}
p = \frac{|j=0 \wedge \Phi_{U} | + |j=1 \wedge \Phi_{\1}|}{\text{all 
cases}}
\end{equation}
\end{enumerate}
\end{scheme}
